---
date: 2017-07-09T13:14:46-05:00
title: "Working with dates and times in R"
summary: ' '
draft: true
---
#Date classes

#Formatting dates

#Storing dates
`POSIXct` is the number of seconds since 12:00 am on January 1st, 1970 (UTC), also known as "unix time." While it may seem ludicrous to represent time this way (as one giant number), this actually has a lot of advantages. You don't have to worry about time zone conversions, and having the time as just a number makes it a lot easier for machines to handle. For human readability, you have to run conversions for day, month, year, and time.

```{r}
as.POSIXct(Sys.Date())
```

`POSIXlt` is a more human-readable form of the time. It is a named list of vectors that includes all of the information you could want about a date.

```{r}
plt <- as.POSIXlt(Sys.Date())
objects(plt)
```

Generally, it's a good idea to store dates or timestamps as `POSIXct` objects. This ensures uniformity, and means that you can operate on all the dates without worrying about differences in how they're stored (month-day-year versus day-month-year, for example). You can always convert these objects to another form if that's what you need.

#Manipulating dates

Dates are just numbers, meaning you can do arithmetic operations with them.

You can add and subtract time:
```{r}

```

You can compare times with logicals:
```{r}

```


#Plotting with dates

They also play well with ggplot, and there are convenient operations you can do specifically on POSIX objects to customize things.